SSA based global code motion of identical computations

### 2017-04-16 acmart-pacmpl-template.tex
barriers

\subsection{Finding candidates to move}
The first step is to find a set of instructions that perform identical
computations: this is performed by a linear scan of all instructions of the
program and classifying all instructions by their value given by \GVN{}. We
could compute available and anticipable sets as computed by GCC's code-hoisting
but that would be a lot of data structures to maintain at each basic block
level. For GCC it makes sense because their code-hoisting is integrated with
GVN-PRE which already has those data structures available.

@@

Code-motion basically consists of two parts i.e., hoisting and sinking. The
unique part about this algorithm is that it only moves identical
computations. An immediate guarantee of this approach is gain in code size (the
final executable may be of larger size because of more inlining). The algorithm
prefers hoisting to sinking. If the dependency of a hoistable candidate is in
the same basic block as the candidate, then the dependency must also be
hoistable otherwise hoisting will be illegal or would require a complicated code
generation to make it legal. The current algorithm discards cases if the
dependency is neither hoistable nor rematerializable.

@@

\section{SSA based global code motion of identical computations}
\label{sec:implementation-details}
% Move this before legality and profitability??
% Ranking of instructions from Zadeck
% Liveness analysis from Ramakrishna

The amount of hoisting depends on whether we collect \GVN{} of instructions
before finding candidates (optimistic) or, on-demand (pessimistic). It also
depends on the generality of the \GVN{} algorithm. We have implemented a
optimistic global code motion of identical computation which uses the liveness
analysis as illustrated in Das \cite{das2012} and ranking expressions explained
in Rosen \cite{rosen1988global}.

\subsection{Optimistic code motion algorithm}
\label{subsec:optimistic}
We collect the \GVN{} of all the instructions in the function and iterate on the
list of instructions having identical \GVN{}s. The algorithm prefers hoisting to
sinking. So first we find the common dominator dominating all such identical
computations and perform legality checks, as described in
Section~\ref{subsec:legality}. Often times it is not possible to hoist all the
instructions to one common dominator, due to legality constraints, e.g.,
intersecting side-effects, or profitability constraints, e.g., increasing
register pressure. In those cases, this algorithmn would partition
(Section~\ref{subsec:partition}) the list of identical instructions into subsets
which can be partially hoisted to their respective common dominators. However,
it should be noted that we only hoist very busy expressions to avoid checking
for undefined behaviors resulting because of introducing extra computation in an
execution path.

computing barriers:

>>>>

\subsection{Finding candidates to move}
\label{subsec:finding-candidates}
The first step is to find a set of instructions that perform identical
computations: this is performed by a linear scan of all instructions of the
program and classifying all instructions by their value given by \GVN{}. We
could compute available and anticipable sets as computed by GCC's code-hoisting
but that would be a lot of data structures to maintain at each basic block
level. For GCC it makes sense because their code-hoisting is integrated with
GVN-PRE which already has those data structures available.

@@

@@

\section{SSA based global code motion of identical computations}
\label{sec:implementation-details}
% Move this before legality and profitability??
% Ranking of instructions from Zadeck
% Liveness analysis from Ramakrishna
In this section we will present an overview of the algorithm we implemented in
\LLVM{}.  The amount of hoisting depends on whether we collect \GVN{} of
instructions before finding candidates (optimistic) or, on-demand
(pessimistic). It also depends on the generality of the \GVN{} algorithm as
mentioned earlier in Section~\ref{subsec:finding-candidates}. We have
implemented a optimistic global code motion of congruent instructions which uses
the liveness analysis as illustrated in Das \cite{das2012} and ranking
expressions explained in Rosen \cite{rosen1988global}.

Code-motion basically consists of two parts i.e., hoisting and sinking. This
implementation only moves congruent instructions. An immediate guarantee of this
approach is gain in code size (the final executable may be of larger size
because of more inlining). The algorithm prefers hoisting to sinking. If the
dependency of a hoistable candidate is in the same basic block as the candidate,
then the dependency must also be hoistable otherwise hoisting will be illegal or
would require a complicated code generation to make it legal. The current
algorithm discards cases if the dependency is neither hoistable nor
rematerializable.

\subsection{Optimistic code motion algorithm}
\label{subsec:optimistic}
We collect the \GVN{} of all the instructions in the function and iterate on the
list of instructions having identical \GVN{}s. The algorithm prefers hoisting to
sinking. So first we find the common dominator dominating all such identical
computations and perform legality checks, as described in
Section~\ref{subsec:legality}. Often times it is not possible to hoist all the
instructions to one common dominator, due to legality constraints, e.g.,
intersecting side-effects, or profitability constraints, e.g., increasing
register pressure; in those cases, this algorithmn would partition
(Section~\ref{subsec:partition}) the list of identical instructions into subsets
which can be partially hoisted to their respective common dominators. However,
it should be noted that we only hoist very busy expressions to avoid checking
for undefined behaviors resulting because of introducing extra computation in an
execution path.

\begin{verbatim}
computing barriers:
for each basic block B in a function:
  barrier_found = false
  for each instruction I in B:
    if I does not guarantee progress:
      mark I as a barrier instruction
      barrier_found = true
      break;

  // Find the last barrier below which instructions can be sunk.
  // If there was no barrier in B, any instruction satisfying
  // other legality checks can be sunk.
  if barrier_found:
    for each instruction I in B in reverse order:
      if I does not guarantee progress:
        mark I as a sink barrier
        break
\end{verbatim}
